---
title: "chi_squared"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(broom)
library(gt)
```


```{r load_data}

# reads in the high rep data from shiny

random_longer <- readRDS("shiny_app/simplified.RDS") 


# I didn't want to change the name of the object read into the ShinyApp
# so I just redid the calculation here

real <- official_housing %>% 
  
  # 1) take the sheet of freshman and nest
  nest(data=everything()) %>% 
  
  # 2) take each tibble of "assigned" freshmen and generate community summaries
  
  # neighborhoods
  mutate(river_central = map(data, ~overview(assigned = ., community = 'river_central'))) %>% 
  mutate(river_east = map(data, ~overview(assigned = ., community = 'river_east'))) %>% 
  mutate(river_west = map(data, ~overview(assigned = ., community = 'river_west'))) %>% 
  mutate(river = map(data, ~overview(assigned = ., community = 'river'))) %>% 
  mutate(quad = map(data, ~overview(assigned = ., community = 'quad'))) %>% 
  
  #river_central
  mutate(adams = map(data, ~overview(assigned = ., community = 'Adams'))) %>% 
  mutate(lowell = map(data, ~overview(assigned = ., community = 'Lowell'))) %>% 
  mutate(quincy = map(data, ~overview(assigned = ., community = 'Quincy'))) %>% 
  
  
  # river_east
  mutate(leverett = map(data, ~overview(assigned = ., community = 'Leverett'))) %>% 
  mutate(dunster = map(data, ~overview(assigned = ., community = 'Dunster'))) %>% 
  mutate(mather = map(data, ~overview(assigned = ., community = 'Mather'))) %>%  
  
  
  # river_west
  mutate(winthrop = map(data, ~overview(assigned = ., community = 'Winthrop'))) %>% 
  mutate(kirkland = map(data, ~overview(assigned = ., community = 'Kirkland'))) %>% 
  mutate(eliot = map(data, ~overview(assigned = ., community = 'Eliot'))) %>%   
  
  
  # quad
  mutate(currier = map(data, ~overview(assigned = ., community = 'Currier'))) %>% 
  mutate(pforzheimer = map(data, ~overview(assigned = ., community = 'Pforzheimer'))) %>% 
  mutate(cabot = map(data, ~overview(assigned = ., community = 'Cabot'))) 

```


```{r random_pivot}

# modifies pfohos name to match in the join

random_longer$community[which(random_longer$community == "pfoho")] <- "pforzheimer"

# pivots the high rep data so that it can be aggregated more easily and ungroups
# replicate

cleaned_random <- random_longer %>%
  ungroup(replicate) %>% 
  select(-replicate) %>% 
  unnest(demographics) %>% 
  select(1:5,8)

aggregated <- aggregate(. ~ community, cleaned_random, mean)


```

```{r real_pivot}

# does the same pivot and selection process for the real data

real_pivoted <- real %>% 
  pivot_longer(-data, 
               names_to = "community", 
               values_to = "demographics") %>% 
  select(-data)


official <- real_pivoted %>%
  unnest(demographics) %>% 
  select(1:5,8)

official
```

```{r join_data}

# two datasets to compare: aggregated (randomized) and official 

aggregated <- aggregated %>% arrange(community)
official <-  official %>% arrange(community)

# joins the aggregated (randomized) and official data to put variables side
# by side for comparison

comparison <- left_join(aggregated, official, by = "community", suffix = c(".aggregated", ".official")) 

community_sizes <- tibble(
  communities = c(
    adams = 132,
    cabot = 126,
    currier = 126,
    dunster = 129,
    eliot = 148,
    kirkland = 130,
    leverett = 156,
    lowell = 150,
    mather = 124,
    pforzheimer = 152,
    quad = 404,
    quincy = 147,
    river = 1233,
    river_central = 429,
    river_east = 409,
    river_west = 395,
    winthrop = 117
  )
)

# changes proportions to frequencies to compare with chi squared
#order is not correct for neighborhoods versus houses


for (i in 2:11){
  comparison[i] <- round(comparison[i]*community_sizes, 0)
}

aggregated
official
comparison

```



```{r split_house_and_neighborhood}

# create house and neighborhood subsets for comparison

comparison_houses <- comparison %>% 
  filter(community %in% c("adams", "cabot", "currier",
                          "dunster", "eliot", "kirkland",
                          "leverett", "lowell", "mather",
                          "pforzheimer", "quincy", "winthrop"))

comparison_neighborhoods <- comparison %>% 
  filter(community %in% c("quad", "river_east", "river_central", "river_west"))

comparison_houses
comparison_neighborhoods

```

```{r manual_method_houses}

# average of (observed - expected)/expected

prop_difference <- tibble(
  prop_international = 0,
  prop_varsity = 0,
  prop_legacy = 0,
  prop_financial_aid = 0,
  prop_group_size = 0
)

for (i in 1:5){
  
  # (observed (real) - expected (random)) / expected (random) for variable i
  
  residual <- abs((comparison_houses[i+6] - comparison_houses[i+1])/comparison_houses[i+1])
  
  
  # take the average residual across all 12 houses for variable i
  
  avg_residual <- sum(residual)/12 
  
  
  # assigns the average residual to prop_difference for variable i
  
  prop_difference[i] = avg_residual

  # repeats the process for all 5 variables
  
}

# Prop_difference represents the proportion difference between each
# variable. Now I take the average for all variables. I use the average
# to get a percent difference that keeps the data consistent in terms of units.

sum(colSums(prop_difference)) / 5 


# avg(|obs - expected| / expected)
# = sum(|obs - expected| / expected)/num_variables
```

```{r manual_method_neighborhoods}

# average of (observed - expected)/expected

prop_difference <- tibble(
  prop_international = 0,
  prop_varsity = 0,
  prop_legacy = 0,
  prop_financial_aid = 0,
  prop_group_size = 0
)

for (i in 1:5){
  
  # (observed (real) - expected (random)) / expected (random) for variable i
  
  residual <- abs((comparison_neighborhoods[i+6] - comparison_neighborhoods[i+1])/comparison_neighborhoods[i+1])
  
  
  # take the average residual across all 12 houses for variable i
  
  avg_residual <- sum(residual)/4
  
  
  # assigns the average residual to prop_difference for variable i
  
  prop_difference[i] = avg_residual

  # repeats the process for all 5 variables
  
}

# Prop_difference represents the proportion difference between each
# variable. Now I take the sum of all the variables.  

sum(colSums(prop_difference)) / 5


# avg(|obs - expected| / expected)
# = sum(|obs - expected| / expected)/num_variables
```





# Results and Concerns:
Is the Chi-squared test appropriate?
- Consider doing it manually without the square. Absolute value difference. What number of houses are more than 4% off (epsilon-delta?)
- The sampling method is simple random sampling (Not true? Collected 600 responses over email; also variables are related to one another (financial aid can be related to ethnicity))
- The variables under study are each categorical (Observations report counts for each categorical variable)
- If sample data are displayed in a contingency table, the expected frequency count for each cell of the table is at least 5 (Yes; enough data)

Interpretation of Chi-squared:
- The p-values are all very high (> 0.20) --> we accept the null hypothesis (we can't reject the null hypothesis) that the data are similar.


modify aggregate to multiply by house size not class

