---
title: "chi_squared"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(broom)
library(gt)
```


```{r load_data}

random_longer <- readRDS("shiny_app/simplified.RDS") 


# I didn't want to change the name of the object read into the ShinyApp
# so I just redid the calculation here

real <- official_housing %>% 
  
  # 1) take the sheet of freshman and nest
  nest(data=everything()) %>% 
  
  # 2) take each tibble of "assigned" freshmen and generate community summaries
  
  # neighborhoods
  mutate(river_central = map(data, ~overview(assigned = ., community = 'river_central'))) %>% 
  mutate(river_east = map(data, ~overview(assigned = ., community = 'river_east'))) %>% 
  mutate(river_west = map(data, ~overview(assigned = ., community = 'river_west'))) %>% 
  mutate(river = map(data, ~overview(assigned = ., community = 'river'))) %>% 
  mutate(quad = map(data, ~overview(assigned = ., community = 'quad'))) %>% 
  
  #river_central
  mutate(adams = map(data, ~overview(assigned = ., community = 'Adams'))) %>% 
  mutate(lowell = map(data, ~overview(assigned = ., community = 'Lowell'))) %>% 
  mutate(quincy = map(data, ~overview(assigned = ., community = 'Quincy'))) %>% 
  
  
  # river_east
  mutate(leverett = map(data, ~overview(assigned = ., community = 'Leverett'))) %>% 
  mutate(dunster = map(data, ~overview(assigned = ., community = 'Dunster'))) %>% 
  mutate(mather = map(data, ~overview(assigned = ., community = 'Mather'))) %>%  
  
  
  # river_west
  mutate(winthrop = map(data, ~overview(assigned = ., community = 'Winthrop'))) %>% 
  mutate(kirkland = map(data, ~overview(assigned = ., community = 'Kirkland'))) %>% 
  mutate(eliot = map(data, ~overview(assigned = ., community = 'Eliot'))) %>%   
  
  
  # quad
  mutate(currier = map(data, ~overview(assigned = ., community = 'Currier'))) %>% 
  mutate(pforzheimer = map(data, ~overview(assigned = ., community = 'Pforzheimer'))) %>% 
  mutate(cabot = map(data, ~overview(assigned = ., community = 'Cabot'))) 

```


```{r}

random_longer$community[which(random_longer$community == "pfoho")] <- "pforzheimer"

cleaned_random <- random_longer %>%
  ungroup(replicate) %>% 
  select(-replicate) %>% 
  unnest(demographics) %>% 
  select(1:5,8)

aggregated <- aggregate(. ~ community, cleaned_random, mean)


```

```{r}
real_pivoted <- real %>% 
  pivot_longer(-data, 
               names_to = "community", 
               values_to = "demographics") %>% 
  select(-data)


official <- real_pivoted %>%
  unnest(demographics) %>% 
  select(1:5,8)

official
```

```{r}

# two datasets to compare: aggregated and official


aggregated
official

comparison <- left_join(aggregated, official, by = "community", suffix = c(".aggregated", ".official")) 

# changes to frequencies to compare with chi squared

comparison[2:11] <- round(comparison[2:11]*1637, 0)

```



```{r}

# create house and neighborhood subsets

comparison_houses <- comparison %>% 
  filter(community %in% c("adams", "cabot", "currier",
                          "dunster", "eliot", "kirkland",
                          "leverett", "lowell", "mather",
                          "pforzheimer", "quincy", "winthrop"))

comparison_neighborhoods <- comparison %>% 
  filter(community %in% c("quad", "river_east", "river_central", "river_west"))

comparison_houses
comparison_neighborhoods

```


```{r}

# compare houses 
tibble(
  variable = c("International", "Varsity", "Legacy", "Financial Aid", "Group Size"),
  pvalue = c(
    # prop_international
    chisq.test(comparison_houses$prop_international.aggregated, comparison_houses$prop_international.official)%>% tidy() %>% pull(p.value),
    
    # prop_varsity
    chisq.test(comparison_houses$prop_varsity.aggregated, comparison_houses$prop_varsity.official) %>% tidy() %>% pull(p.value),
    
    # prop_legacy
    chisq.test(comparison_houses$prop_legacy.aggregated, comparison_houses$prop_legacy.official) %>% tidy() %>% pull(p.value),
    
    # prop_financial_aid
    chisq.test(comparison_houses$prop_financial_aid.aggregated, comparison_houses$prop_financial_aid.official) %>% tidy() %>% pull(p.value),
    
    # prop_group_size
    chisq.test(comparison_houses$prop_group_size.aggregated, comparison_houses$prop_group_size.official) %>% tidy() %>% pull(p.value)
  )
) %>% 
  gt() %>% 
  tab_header(
    title = "Chi-squared test ",
    subtitle = "Comparison between the twelve Harvard houses"
  ) %>% 
  cols_label(
    variable = "Variable",
    pvalue = "P-Value"
  )
```

```{r}
# compare neighborhoods 

tibble(
  variable = c("International", "Varsity", "Legacy", "Financial Aid", "Group Size"),
  pvalue = c(
    # prop_international
    chisq.test(comparison_neighborhoods$prop_international.aggregated, comparison_neighborhoods$prop_international.official) %>% tidy() %>% pull(p.value),
    
    # prop_varsity
    chisq.test(comparison_neighborhoods$prop_varsity.aggregated, comparison_neighborhoods$prop_varsity.official) %>% tidy() %>% pull(p.value),
    
    # prop_legacy
    chisq.test(comparison_neighborhoods$prop_legacy.aggregated, comparison_neighborhoods$prop_legacy.official) %>% tidy() %>% pull(p.value),
    
    # prop_financial_aid
    chisq.test(comparison_neighborhoods$prop_financial_aid.aggregated, comparison_neighborhoods$prop_financial_aid.official) %>% tidy() %>% pull(p.value),
    
    # prop_group_size
    chisq.test(comparison_neighborhoods$prop_group_size.aggregated, comparison_neighborhoods$prop_group_size.official) %>% tidy() %>% pull(p.value)
  )
) %>% 
  gt() %>% 
  tab_header(
    title = "Chi-squared test ",
    subtitle = "Comparison between the four Harvard neighborhoods"
  ) %>% 
  cols_label(
    variable = "Variable",
    pvalue = "P-Value"
  )

    
```

# Results and Concerns:
Is the Chi-squared test appropriate?
- The sampling method is simple random sampling (Not true? Collected 600 responses over email; also variables are related to one another (financial aid can be related to ethnicity))
- The variables under study are each categorical (Observations report counts for each categorical variable)
- If sample data are displayed in a contingency table, the expected frequency count for each cell of the table is at least 5 (Yes; enough data)

Interpretation of Chi-squared:
- The p-values are all very high (> 0.20) --> we accept the null hypothesis (we can't reject the null hypothesis) that the data are similar.

